Analysis

### 2023-03-29 Analysis.rmd
Reorganise code that checks for the number of participants, add plot about expert participation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

colors <- RColorBrewer::brewer.pal(n = 9, name = "Set1")

@@

@@

```{r}
# helper function to score forecasts both on the natural and on the log scale
score_and_add_log <- function(forecasts) {
  forecasts |>
  mutate(scale = "natural") |>
  rbind(forecasts |>
          mutate(
            scale = "log", 
            true_value = log(true_value + 1), 
            prediction = log(pmax(prediction, 0) + 1)
          )) |>
  score(metrics = c("interval_score", "coverage")) |>
  add_coverage(by = c("target_type", "horizon", "method")) |>
    add_type_and_scale() |>
  as.data.table() 
}

# helper function that filters the forecast data such that we only have 
# forecasts left where a single person made a forecast both for direct and rt
filter_comparison_rt_direct <- function(forecasts) {
  forecasts_comparison <- forecasts |>
  filter(model %in% models_human) |>
  mutate(rt_name = ifelse(
    grepl("(Rt)", model), 
    gsub("\ \\(Rt\\)", "", model), 
    NA
  )) |>
  mutate(method = ifelse(grepl("(Rt)", model), "rt", "direct")) |>
  group_by(forecast_date, target_type) |>
  filter(model %in% rt_name | rt_name %in% model) |>
  group_by(forecast_date, target_type, method) |>
  select(target_end_date, target_type, forecast_date, horizon, model, method, true_value, prediction, quantile)
}

@@

## Number of forecasts and ensemble members
avail <- 
  available(forecasts, filter = models_direct, name = "Direct") |>
  rbind(available(forecasts, filter = models_rt, name = "Rt")) |>
  mutate(group = "Human forecast")

@@

plot_number_forecasters <- avail |>
  pivot_wider(names_from = type, values_from = `Number forecasts`) |>
  mutate(Rt = Direct + Rt) |>
  pivot_longer(cols = c(Direct, Rt), names_to = "type",
               values_to = "Number forecasts") |>
  as.data.table() |>
  rbind(avail_ensemble, fill = TRUE) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             group = group,
             fill = group, 
             alpha = type)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

@@

avail |> 
  group_by(target_type, type) |>
  summarise(
    mean = mean(`Number forecasts`), 
    median = median(`Number forecasts`), 
    min = min(`Number forecasts`), 
    max = max(`Number forecasts`)
  )

## number of unique participants
unique_participants <- gsub("\ \\(Rt\\)", "", models_human) |> unique()
length(unique_participants)

## Number of unique participants in every week
forecasts |>
  filter(model %in% models_human) |>
  mutate(model = gsub("\ \\(Rt\\)", "", model)) |>
  select(model, forecast_date) |>
  group_by(forecast_date) |>
  summarise(n = length(unique(model))) |>
  summary()

## number of times a forecaster submitted
forecasts |>
  filter(model %in% models_human) |>
  mutate(model = gsub("\ \\(Rt\\)", "", model)) |>
  select(forecast_date, model) |>
  group_by(model) |>
  summarise(n = length(unique(forecast_date))) |>
  summary()
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)
library(scales)

colors <- RColorBrewer::brewer.pal(n = 9, name = "Set1")
colors <- hex <- hue_pal()(4)

@@

# helper function to score forecasts both on the natural and on the log scale
score_and_add_log <- function(forecasts) {
  forecasts |>
  mutate(scale = "natural") |>
  rbind(forecasts |>
          mutate(
            scale = "log", 
            true_value = log(true_value + 1), 
            prediction = log(pmax(prediction, 0) + 1)
          )) |>
  score(metrics = c("interval_score", "coverage")) |>
  add_coverage(by = c("target_type", "horizon", "method")) |>
    add_type_and_scale() |>
  as.data.table() 
}

@@

```{r}
# helper function that filters the forecast data such that we only have 
# forecasts left where a single person made a forecast both for direct and rt
filter_comparison_rt_direct <- function(forecasts) {
  forecasts_comparison <- forecasts |>
  filter(model %in% models_human) |>
  mutate(rt_name = ifelse(
    grepl("(Rt)", model), 
    gsub("\ \\(Rt\\)", "", model), 
    NA
  )) |>
  mutate(method = ifelse(grepl("(Rt)", model), "rt", "direct")) |>
  group_by(forecast_date, target_type) |>
  filter(model %in% rt_name | rt_name %in% model) |>
  group_by(forecast_date, target_type, method) |>
  select(target_end_date, target_type, forecast_date, horizon, model, method, true_value, prediction, quantile)
}

@@

## Number of forecasts 
# this gives the total with people being double counted if they submit both
# Rt and direct forecasts
avail <- 
  available(forecasts, filter = models_direct, name = "Direct") |>
  rbind(available(forecasts, filter = models_rt, name = "Rt")) |>
  mutate(group = "Human forecast")

@@

plot_number_forecasters <- avail |>
  pivot_wider(names_from = type, values_from = `Number forecasts`) |>
  mutate(Rt = Direct + Rt) |>
  pivot_longer(cols = c(Direct, Rt), names_to = "type",
               values_to = "Number forecasts") |>
  as.data.table() |>
  rbind(avail_ensemble, fill = TRUE) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             group = group,
             fill = group, 
             alpha = type)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  scale_fill_manual(values = colors[1:2]) +
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

@@

## overall number of unique participants
unique_participants <- gsub("\ \\(Rt\\)", "", models_human) |> unique()
length(unique_participants)

## overall number of experts
forecasts |>
  filter(model %in% models_human) |>
  update_expert_status() |>
  mutate(model = gsub("\ \\(Rt\\)", "", model)) |>
  select(model, expert) |>
  unique() |>
  group_by(expert) |>
  summarise(n = length(unique(model)))

## number of times a forecaster submitted
forecasts |>
  filter(model %in% models_human) |>
  mutate(model = gsub("\ \\(Rt\\)", "", model)) |>
  select(forecast_date, model) |>
  group_by(model) |>
  summarise(n = length(unique(forecast_date))) |>
  summary()

# helper function
sum_number <- function(data) {
  data |> 
    mutate(n = n()) |>
    summarise(
      mean = mean(`n`), 
      median = median(`n`), 
      min = min(`n`), 
      max = max(`n`)
    )
}

nums <- forecasts |>
  update_expert_status() |>
  filter(model %in% models_human) |>
  mutate(method = ifelse(grepl("(Rt)", model), "Rt", "direct")) |>
  mutate(model = gsub("\ \\(Rt\\)", "", model)) |>
  select(model, forecast_date, target_type, expert, method) |>
  unique()

# number of forecasts, (double counting participants who made a forecast for both Rt and direct)
nums |>
  group_by(target_type, forecast_date, method) |>
  summarise(n = n()) |>
  group_by(target_type, method) |>
  summarise(
    mean = mean(`n`), 
    median = median(`n`), 
    min = min(`n`), 
    max = max(`n`)
  )

# unqiue forecasts for cases and deaths without double counting
nums |>
  select(-method) |>
  unique() |>
  group_by(forecast_date, target_type) |>
  summarise(n = n()) |>
  group_by(target_type) |>
  summarise(
    mean = mean(`n`), 
    median = median(`n`), 
    min = min(`n`), 
    max = max(`n`)
  )

# unique participants every week (without double counting Rt and direct)
nums |>
  select(-method, -target_type) |>
  unique() |>
  group_by(forecast_date) |>
  summarise(n = n()) |>
  summarise(
    mean = mean(`n`), 
    median = median(`n`), 
    min = min(`n`), 
    max = max(`n`)
  )

# number of forecasts from experts and non-experts every week (allowing double counting)
nums |>
  group_by(forecast_date, target_type, expert) |>
  summarise(n = n()) |>
  group_by(target_type, expert) |>
  summarise(
    mean = mean(`n`), 
    median = median(`n`), 
    min = min(`n`), 
    max = max(`n`)
  )

p_num_experts <- nums |>
  mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
  group_by(target_type, forecast_date, expert, method) |>
  summarise(n = n()) |>
  pivot_wider(names_from = method, values_from = n) |>
  mutate(Rt = Rt + direct) |>
  pivot_longer(cols = c(Rt, direct), names_to = "method", values_to = "n") |>
  ggplot(aes(x = forecast_date, y = n, 
             group = expert,
             fill = expert, 
             alpha = method)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

ggsave("output/figures/num-forecasters-expert.png", plot = p_num_experts, 
       width = 7, height = 3.5)

# forecasts |> 
#   update_expert_status() |>
#   mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
#   filter(forecast_date == "2021-06-07", 
#          model %in% models_human, horizon == 1, 
#          quantile == 0.5) |> 
#   mutate(method = ifelse(grepl("(Rt)", model), "Rt", "direct")) |>
#   mutate(model = gsub("\ \\(Rt\\)", "", model)) |> 
#   group_by(target_type, method, expert) |> 
#   summarise(n = length(unique(model)))
